---
title: "Useful Plots"
output: html_notebook
---
 
#### Use the forecast library
```{r}
library("forecast")
```

#### Read in the Amtrack Data
```{r}
Amtrak.data <- read.csv("Amtrak data(1).csv")
```

#### Fit Time Series Model to our Data
```{r}
ridership.ts <- ts(Amtrak.data$Ridership, start = c(1991,1), end = c(2004, 3), freq = 12)
```

#### Time Series Plot
```{r}
plot(ridership.ts, xlab = "Time", ylab = "Ridership", ylim = c(1300, 2300), bty = "l")
```

#### Seasonal Plot
```{r}
seasonplot(ridership.ts, ylab="Ridership", 
           xlab="Year", main="Seasonal Plot", year.labels=TRUE)
```
Plots each month by year. We can see there's a definite pattern in our data. WE have low ridership in Jan and Feb. Ridership increases through the summer months. Ridership dips down in september and (usually) recovers by Dec. 

#### Seasonal Deviation Plot
```{r}
monthplot(ridership.ts, ylab="Ridership", 
           xlab="Year", main="Seasonal Deviation Plot")
```
Shows us the trend
Example: The first "J" shows January versus January
We see a repeated trend; for the first few months we see a dip, followed by a recovery. Towards the summer months we just see recovery.  

#### Lag Plot for our Time Series
```{r}
lag.plot(ridership.ts, lags=16)
```
Y versus Y Lag 1, Y Lag 2, etc.  
We can inspect for linear relationships. Notice the plots for Lag 1 and Lag 12. 

#### Time Series Plot with Auto and Partial Correlation Graphs
```{r}
tsdisplay(ridership.ts)
```
In one plot we can look at our

* Time Series Plot  
* Auto Correlations  
* Partial Auto Correlations
We can see the Partial Correlation spikes at 1,6(neg), 7  10, 11, 12, 13(neg). Some of these patterns. don't really make any sense. 






