---
title: "time_series_graphics_ex"
author: "Tommy Englert"
date: "November 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(forecast)
library(fpp2)
library(tidyverse)
```

## EXERCISE 1  

```{r}
#help(gold)
frequency(gold)

```

```{r}
autoplot(gold)
```

Daily morning gold prices in US dollars. 
1 January 1985 - 31 March 1989.

What is the outlier in our gold time series?
```{r}
which.max(gold)
```






```{r}
#help('woolyrnq')
frequency(woolyrnq)
```

```{r}
autoplot(woolyrnq)
```

Quarterly production of woollen yarn in Australia: tonnes. 
Mar 1965 - Sep 1994.

```{r}
#help(gas)
frequency(gas)

```

```{r}
autoplot(gas)
```

Australian monthly gas production: 1956-1995.

## EXERCISE 2  

```{r}
tute1 <- read.csv("tute1.csv")
```
Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product.

#### Convert the data to a time series  

```{r}
str(tute1)
```

```{r}
mytimeseries <- ts(tute1[,-1], start = 1981, frequency = 4)
# [,-1] removes the first column from the time series model because we don't need the quarters
```

#### Construct time series plots for each of the series

```{r}
autoplot(mytimeseries, facets = TRUE)
```


## EXERCISE 3  

```{r}
retaildata <- readxl::read_excel("retail.xlsx", skip = 1)
```

chose one of the 40 columns to explore
```{r}
myts <- ts(retaildata[,"A3349902A"], frequency=12, start=c(1982,4))
#observations are on the first of every month
```

#### Explore your chosen time series using the following functions:  

* autoplot  
* ggseasonplot  
* ggsubseriesplot  
* gglagplot  
* ggAcf

Can you spot any seasonality, cyclicity and trend? 
What do you learn about the series?

```{r}
autoplot(myts)
```

```{r}
ggseasonplot(myts)
```


```{r}
ggseasonplot(myts, polar = TRUE)
```


```{r}
ggsubseriesplot(myts)
```

```{r}
gglagplot(myts)
```

```{r}
ggAcf(myts)
```


## EXERCISE 4  
```{r}
help(bicoal)
```

```{r}
tsdisplay(bicoal)
```


```{r}
autoplot(bicoal)
#ggseasonplot(bicoal)
#ggseasonplot(bicoal, polar = TRUE)
ggsubseriesplot(bicoal)
gglagplot(bicoal)
ggAcf(bicoal)


```

```{r}
autoplot(chicken)
#ggseasonplot(chicken) <- annual data means no season plot
#ggseasonplot(bicoal, polar = TRUE) <- annual data means no season plot
#ggsubseriesplot(chicken) <- annual data means no subseries plot
gglagplot(chicken)
ggAcf(chicken)
```

```{r}
autoplot(dole)
ggseasonplot(dole)
ggseasonplot(dole, polar = TRUE)
ggsubseriesplot(dole)
gglagplot(dole)
ggAcf(dole)

```

```{r}
str(usdeaths)
```

```{r}
autoplot(usdeaths)
ggseasonplot(usdeaths)
ggseasonplot(usdeaths, polar = TRUE)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths)
```





















































